---
title: "A simple Classifier"
author: "A. Adriano"
date: "Statistical Learning"
output: html_document
---
<style type="text/css">

h1.title {
  font-size: 42px;
  color: DarkRed;
  text-align: center;
}
h4.author { //
    font-size: 25px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { //
  font-size: 20px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
</style>
****
### A Simple Classifier on Iris Data
</br>

First, I import the data.
```{r echo=FALSE, message=FALSE, warning= FALSE}
library(tidyverse)
library(kableExtra)
iris <- as_tibble(read_csv('iris.csv'))
```


```{r}
(glimpse(iris))
```
</br>
Next, I need to partition the data into the test and training set (80% & 20%, respectively).

```{r message=FALSE, warning= FALSE}
# Partition the data
#install.packages("caret")
library(caret)
partition <- createDataPartition(iris$species, p=0.80, list=FALSE)
# 80% to train, 20% to test
test <- iris[-partition,]
train <- iris[partition,]
```

I plot the data to get a general idea of the characteristics and the distribution:
```{r}
iris %>%
  ggplot(aes(sepal_length, sepal_width, color=species, shape=species)) +
    geom_point ()+
    xlab("Petal Length") +
    ylab("Petal Width") +
    ggtitle("Petal Length-Width")
```

Next, I prepare my algorithm for the control and fit a simple linear classifier onto the training data.
```{r}
library(pander)
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
#Using a simple linear algorithm:
set.seed(7)
fit.lda <- train(species~., data=train, method="lda", metric=metric, trControl=control)
#See the outcomes
pander(print(fit.lda))
```

The next step is to estimate the model onto the remaining test set. This is the outcome for recognizing all 3 flower types simultaneously: "I. Setosa"" versus "I. Versicolor"" versus "I. Virginica".
Predictions:
```{r}
# estimate skill of LDA on the test set
skill <- predict(fit.lda, test)
skill
```
Actual:
```{r}
test$species
```
```{r}
#confusionMatrix(skill, test$species)
```

Confusion Matrix and Statistics
Reference

|Prediction | Setosa | Versicolor | Virginica |
| ------------- |:-------------:| -----:| -----:|
| Setosa      | 10 | 0 |  0 |
| Versicolor    | 0     |   10 |  0 |
| Virginicas | 0    |    0 | 10 |
 
Overall Statistics
                                     
               Accuracy : 1          
                 95% CI : (0.8843, 1)
    No Information Rate : 0.3333     
    P-Value [Acc > NIR] : 4.857e-15  
                                     
                  Kappa : 1          
 
|Statistics by Class:| Class: setosa | Class: versicolor | Class: virginica |
| ------------- |:-------------:| -----:| -----:|
|Sensitivity            |     1.0000  |          1.0000 |          1.0000 |
|Specificity              |   1.0000   |         1.0000  |         1.0000|
|Pos Pred Value            |  1.0000    |        1.0000   |        1.0000|
|Neg Pred Value             | 1.0000     |       1.0000    |       1.0000 |
|Prevalence       |           0.3333      |      0.3333     |      0.3333|
|Detection Rate    |          0.3333       |     0.3333      |     0.3333|
|Detection Prevalence |       0.3333        |    0.3333       |    0.3333|
|Balanced Accuracy     |      1.0000         |   1.0000        |   1.0000|
