---
title: "Perceptron Problem"
author: "A. Adriano"
date: "Statistical Learning"
output: html_document
---
<style type="text/css">

h1.title {
  font-size: 42px;
  color: DarkRed;
  text-align: center;
}
h4.author { //
    font-size: 25px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { //
  font-size: 20px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
</style>
****
### Perceptron algorithm for the Iris Dataset
</br>

First, I import the data. I was asked to write code for two dimensions, so I chose length and width. I apply it to the iris data from Problem 1 as an example.
```{r echo=FALSE, message=FALSE, warning= FALSE}
library(tidyverse)
library(kableExtra)
iris <- as_tibble(read_csv('iris.csv'))
```

```{r}
(glimpse(iris))
```

Next, the Perceptron will work better using binary data, so let Setosa = -1 and let all others by = 1.
```{r}
iris[, 4] <- 1
iris[iris[, 3] == "setosa", 4] <- -1

x <- iris[, c(1, 2)]
y <- iris[, 4]

# head and tail of data 
head(x)
head(y)
```
Next I need to write the algorithm that can take the data and number of iterations through the program and update the weights of the factors.
```{r}
perceptron <- function(x, y, eta, niter) {
        #Start with a weight vector
        weight <- rep(0, dim(x)[2] + 1)
        errors <- rep(0, niter)
        # create a loop off the training set
        for (jj in 1:niter) {
                for (ii in 1:length(y)) {
                        #Make the prediciton
                        z <- sum(weight[2:length(weight)] * 
                                         as.numeric(x[ii, ])) + weight[1]
                        if(z < 0) {
                                ypred <- -1
                        } else {
                                ypred <- 1}
                        # adjust the weight
                        weight2 <- eta * (y[ii] - ypred) * 
                                c(1, as.numeric(x[ii, ]))
                        weight <- weight + weight2
                        # adjust error
                        if ((y[ii] - ypred) != 0.0) {
                                errors[jj] <- errors[jj] + 1
                        }}}
# weight to decide between the two species 
print(weight)
return(errors)
}
```

Next, we extend to all 3 types of flowers,
```{r}
iris2 <- iris[, c(1, 3, 5)]
names(iris2) <- c("sepal", "petal", "species")
```
And print the error:
```{r}
x <- iris[, 1:4] 
names(x) <- tolower(names(x))
y <- rep(-1, dim(x)[1])
y[iris[, 5] == "virginica"] <- 1


err <- perceptron(x, y, 0.01, 50)
```

